metric-learn: Metric Learning Algorithms in Python

### 2019-08-07 metric-learn-jmlr.tex
Add back affiliation
Updates from Overleaf

\usepackage{jmlr2e}
\usepackage{minted}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{stmaryrd}
\usepackage[colorinlistoftodos]{todonotes}

@@

\ShortHeadings{Metric-learn: Metric Learning in Python}{de Vazelhes, Carey, Tang, Vauquier and Bellet}
\firstpageno{1}

@@

\title{Metric-learn: Metric Learning in Python}

\author{\name William de Vazelhes \email william.de-vazelhes@inria.fr \\
       \addr INRIA, France
       \AND
       \name CJ Carey \email perimosocordiae@gmail.com \\
       \addr \todo[inline]{TODO}
       \AND
       \name Yuan Tang \email terrytangyuan@gmail.com \\
       \addr \todo[inline]{TODO}
       \AND
       \name Nathalie Vauquier \email nathalie.vauquier@inria.fr \\
       \addr INRIA, France
       \AND
       \name Aur\'elien Bellet \email aurelien.bellet@inria.fr \\
       \addr INRIA, France
       }

@@

\begin{abstract}%   <- trailing '%' for backward compatibility of .sty file
\texttt{metric-learn} is an open source Python package for metric-learning, based on \texttt{numpy}, \texttt{scipy}, and \texttt{scikit-learn}. It allows to learn a metric between data points that satisfy desired properties. \texttt{metric-learn} stands out from other packages by providing a unified interface for metric learning with several modes of supervision (full supervision or pairs and quadruplets supervision), all being compatible with \texttt{scikit-learn} and allowing to do cross-validation, model selection, and pipelining. The package is released under the MIT licence, and the current version (v0.5.0) is downloadable from PyPi. The source code is available online at \url{http://github.com/metric-learn/metric-learn}, as well as the documentation at \url{http://metric-learn.github.io/metric-learn/}.
\end{abstract}

@@

Many approaches in machine learning require a measure of distance between data
points. Traditionally, practitioners would choose a standard distance metric
(Euclidean, City-Block, Cosine, etc.) using a priori knowledge of the
domain. However, it is often difficult to design metrics that are well-suited
to the particular data and task of interest.
% \todo[inline]{probably need to add a sentence explaining what is meant by weak supervision}
Distance metric learning  (or simply, metric learning) (\cite{Bellet13}, \cite{Kulis13}, \cite{Surez18}) aims at
automatically constructing task-specific distance metrics from data, in a machine learning manner. It can deal with various types of supervision: either full supervision, where we know label(s) for each sample, or weaker supervision, where we only have some information about the similarities between samples. The learned distance metric can
then be used to perform various tasks, for instance to retrieve the elements (images, documents) of a database that are semantically closer to a query element. It can also be plugged into other machine learning algorithms: popular use-cases include improving the accuracy of nearest neighbors learning models (for classification, regression, anomaly detection...) or biasing the clusters found by clustering algorithms like K-Means towards the intended semantics. Finally, metric learning can be used to perform dimensionality reduction.
These use-cases highlight the importance of integrating metric learning into the rest of the machine learning pipeline and tools.
% \todo[inline]{quick presentation of the package: implements several algorithms, with sklearn-compatible API for model selection, model evaluation, pipelining with other estimators. makes metric-learn stands out from other packages (maybe mention a few of them), in particular in the case of weakly supervised algorithms}
\texttt{metric-learn} is a package for metric learning in Python, which provides many common metric-learning algorithms unified through a common interface. It allows to do metric-learning with different level of supervision.
% (not only class supervision, unlike most packages are focused on, including another package for metric learning in Python, \texttt{PyDML})
% , which is one interesting feature allowed by the metric learning paradigm. 
Its API is compatible with scikit-learn, the leading library for machine learning in Python, allowing for model selection, evaluation, and pipelining, which makes \texttt{metric-learn} stand out from other packages, like \texttt{PyDML} \footnote{\url{https://github.com/jlsuarezdiaz/pyDML}}, which mostly focuses on supervised metric learning algorithms.
% Acknowledgements should go at the end, before appendices and references
The rest of this paper is organized as follows. We first present an overview of the package, followed by a more detailed description of the available algorithms. We then give more details about the API and the software architecture, in particular the compatibility with \texttt{scikit-learn}, and conclude with a discussion of future developments.

\texttt{metric-learn} only relies on core libraries from the SciPy ecosystem: \texttt{numpy}, \texttt{scipy}, and \texttt{scikit-learn}.\footnote{An optional dependency with \texttt{skggm} (\cite{skggm}) can be installed in order to use a more stable solver when using the LSML algorithm from the package.}
The current release (v.0.5.0) can be installed from the Python Package Index (PyPI), for python 2.7 and 3.5 or later. The source code is available on GitHub at \url{http://github.com/metric-learn/metric-learn}. The development is collaborative and open to all contributors through the usual GitHub framework (issues and pull requests). The code is free of use, provided under the MIT licence. The interest of the package from the community is clear, since the project has been starred more than 730 times and forked more than 170 times at the time of writing.
A detailed documentation, including installation guidelines, the description of the algorithms and the API, as well as example of use done with \texttt{sphinx-gallery}, is available at \url{https://metric-learn.github.io/metric-learn/}. The quality of the code is ensured by an intensive test coverage of 96\%. Every new contribution to the code is automatically checked by a continuous integration platform that ensures a minimal test coverage of the added code, and an minimal variation in test coverage of the overall code, which allows easy contribution for incoming developers, which can have their code automatically checked.

\section{Available algorithms}
% \todo[inline]{comparison with other packages ? }
% Metric learning problems fall into two main categories depending on the type
% of supervision available about the training data:

@@

\paragraph{General setting} Metric learning problems are generally formulated as an optimization problem where one seeks to find, on the training data $(x_0, ..., x_n)$, the parameters $\theta$ of a distance function $D_\theta(x_i, x_j)$ that optimize some desired objective function $\mathcal{L}_\theta(x_0, ..., x_n) = f(D_\theta(x_0, x_0), ..., D_\theta(x_0, x_n), ..., D_\theta(x_n, x_0), ..., D_\theta(x_n, x_n))$, hoping that this metric will generalize the desired propertie(s) on new incoming test data.

\subparagraph{Mahalanobis metric learning}

In the metric-learn package, all algorithms currently implemented learn so-called Mahalanobis distances. Given a real-valued parameter matrix $L$ of shape \texttt{(num\_dims, n\_features)} where \texttt{n\_features} is the
number of features describing the data, the Mahalanobis distance associated with
$L$ is defined as follows:

$D(x, x') = \sqrt{(Lx-Lx')^\top(Lx-Lx')}$

In other words, a Mahalanobis distance is a Euclidean distance after a
linear transformation of the feature space defined by $L$ (taking
$L$ to be the identity matrix recovers the standard Euclidean distance).
Mahalanobis distance metric learning can thus be seen as learning a new
embedding space of dimension \texttt{num\_dims}. Note that when \texttt{num\_dims} is
smaller than \texttt{n\_features}, this achieves dimensionality reduction.

@@

\paragraph{Supervised Learners}
The first possibility to learn this metric is from class supervision. We can learn a metric that brings closer together points from the same class and spreads further away points from a different class. For instance, in figure \ref{fig:full}, the class is the identity of a person of an image, and we want to put closer together images from the same person.Currently, \texttt{metric-learn} contains the following algorithms for full supervision: Neighborhood Components Analysis (NCA) (\cite{Goldberger04}), Large Margin Nearest Neighbors (LMNN) (\cite{Weinberger09}), Relative Components Analysis (RCA) (\cite{Shental02}), Local Fisher Discriminant Analysis (LFDA) (\cite{Sugiyama07}) and Metric Learning for Kernel Regression (MLKR) (\cite{Weinberger07}) (which uses a continuous label like in regressions, instead of a class label).

% \todo[inline]{ces figures ne sont pas très vendeuses. je propose de faire une figure conceptuelle en prenant l'exemple des faces expliquant la différence entre le supervisé, les paires et les quadruplets. on peut s'inspirer de \url{http://cedric.cnam.fr/~thomen/papers/Law_ICCV_2013_QWise.pdf}}

% \begin{figure}[H]
% \centering
% \begin{subfigure}{.5\textwidth}
%   \centering
%   \includegraphics[width=.8\linewidth]{supervised_without_metric.png}
%   \caption{Original points, before metric learning}
%   \label{fig:sub1}
% \end{subfigure}%
% \begin{subfigure}{.5\textwidth}
%   \centering
%   \includegraphics[width=.8\linewidth]{supervised_with_metric.png}
%   \caption{points after metric learning (ITML)}
%   \label{fig:sub2}
% \end{subfigure}
% \caption{Metric learning with pairwise constraints}
% \label{fig:test}
% \end{figure}

% \paragraph{Weakly Supervised Metric Learning algorithms}
% Weakly Supervised algorithms use less information than the previous algorithms: they only take tuple of points and possibly a corresponding label. See below for more concrete examples of algorithms learning on tuples.
\subparagraph{Pairs Learners} 
Another way is to have pair of points as inputs, and a corresponding label indicating whether the two points are similar or not. We can then learn a metric that brings points from a similar pair closer together, and points from a dissimilar pair further away from each other. For instance, we could want to put closer together in the embedding space images from the same person (i.e. similar), and further away images from a different person (i.e. dissimilar) (see figure \ref{fig:pairs}). Currently, \texttt{metric-learn} contains the following algorithms for pairs supervision: Mahalanobis Metric for Clustering (MMC) (\cite{Xiang08}), Sparse High-Dimensional Metric Learning (SDML) (\cite{Qi09}), Information Theoretic Metric Learning (ITML) (\cite{Davis07}).

% \begin{figure}[H]
% \centering
% \begin{subfigure}{.5\textwidth}
%   \centering
%   \includegraphics[width=.8\linewidth]{pairs_without_metric.png}
%   \caption{Original points, before metric learning}
%   \label{fig:sub1}
% \end{subfigure}%
% \begin{subfigure}{.5\textwidth}
%   \centering
%   \includegraphics[width=.8\linewidth]{pairs_with_metric.png}
%   \caption{points after metric learning (ITML)}
%   \label{fig:sub2}
% \end{subfigure}
% \caption{Metric learning with pairwise constraints}
% \label{fig:test}
% \end{figure}

\subparagraph{Quadruplets Learners}
The last possibility to learn a metric in \texttt{metric-learn} currently is to take as input quadruplets of points. We will learn a metric that brings the two first points of each quadruplet closer than the two last points are. This can be used for instance to learn a metric space in which closer points have closer values of a given attributes. In figure \ref{fig:quadruplets} for instance, we want to learn a metric that puts closer together people from approximately the same age. Currently the only algorithm in \texttt{metric-learn} for learning on quadruplets is Metric Learning from Relative Comparisons by Minimizing Squared Residual (LSML) (\cite{Liu12}).

\begin{figure}[H]
  \centering
  \begin{subfigure}[t]{0.3\textwidth}
     \centering 
     \begin{tabular}{c|c}
     \texttt{X} & \texttt{y} \\ \hline \\
   \includegraphics[scale=0.15]{Tom_Cruise_0010.jpg} & 'Tom' \\
   \includegraphics[scale=0.15]{Kate_Winslet_0001.jpg} & 'Kate' \\
\end{tabular}
     \caption{}
     \label{fig:full}
     \end{subfigure}
  \hspace{5pt}
  \begin{subfigure}[t]{0.3\textwidth}
     \centering
     \begin{tabular}{c|c}
     \texttt{pairs} & \texttt{y\_pairs} \\  \hline \\
\includegraphics[scale=0.15]{Aaron_Peirsol_0001.jpg} \includegraphics[scale=0.15]{Aaron_Peirsol_0002.jpg} & 1\\  \includegraphics[scale=0.15]{Angelina_Jolie_0001.jpg}   \includegraphics[scale=0.15]{Kate_Winslet_0001.jpg} & -1 \\
\end{tabular}
     \caption{}
     \label{fig:pairs}
  \end{subfigure}
  \hspace{5pt}
  \begin{subfigure}[t]{0.3\textwidth}
     \centering

          \begin{tabular}{c}
     \texttt{quadruplets} \\ \hline \\
     \includegraphics[scale=0.15]{Kate_Winslet_0001.jpg} \includegraphics[scale=0.15]{Nicole_Kidman_0001.jpg} \includegraphics[scale=0.15]{Daniel_Radcliffe_0001.jpg} \includegraphics[scale=0.15]{Kirk_Douglas_0001.jpg} \\
          \includegraphics[scale=0.15]{Leonardo_DiCaprio_0001.jpg} \includegraphics[scale=0.15]{Angelina_Jolie_0001.jpg} \includegraphics[scale=0.15]{Daryl_Sabara_0001.jpg} \includegraphics[scale=0.15]{Carl_Reiner_0002.jpg} \\
\end{tabular}

     \caption{}
     \label{fig:quadruplets}
  \end{subfigure}
  \caption{Different types of supervision. \textbf{\ref{fig:full} Full supervision:} we want samples from the same label to be closer than samples from different labels. \textbf{\ref{fig:pairs} Pairs supervision:} we want similar samples (1) (the same person) to be closer than dissimilar samples (-1) (different persons). \textbf{\ref{fig:quadruplets} Quadruplets supervision:} no labels. We want the two leftmost samples (not much age difference) to be closer than the two rightmost samples (bigger age difference).}
  \label{fig:contour}
\end{figure}

@@

\texttt{metric-learn}'s goal is to provide a unified interface to all metric-learning algorithms, including the ones that have an input format that differs from classical machine learning ones, and allow to combine them with scikit-learn objects. As such, all metric-learners from \texttt{metric-learn} inherit from an abstract \texttt{BaseMetricLearner} object.
% \todo[inline]{Should I talk about the classes etc ? Not useful since they are not visible by the user right ?}
% \todo[inline]{I think it can be nice, this is not documentation, but explaining how things are coded}

\texttt{metric-learn}'s API is designed to be fully compatible with \texttt{scikit-learn} API (\cite{scikit-learn}).

@@

Supervised metric learners inherit from scikit-learn base class \texttt{TransformerMixin}, just like \texttt{scikit-learn}'s \texttt{LinearDiscriminantAnalysis} for instance, and they behave as such.
% (they pass all scikit-learn's "estimators checks"). 
%  They can easily be pipelined with nearest neighbors estimators to improve their performance: \texttt{better\_knn = sklearn.pipeline.make\_pipeline(NCA(), KNeighborsClassifier())} (TODO: sth to explain NCA is in metric-learn)

\paragraph{Weakly Supervised metric learners} 
% Algorithms that learn on pairs or quadruplets take as input 3D array of shape \texttt{(n\_samples, t, n\_features)} (where...). In other words, sampling one element along the first dimension returns one tuple (e.g. a pair), that is a matrix of t lines corresponding to the t elements in the tuple.
% This format is what allows all these API choices, since slicing along the 1st dimension
% While a classical input array \texttt{X} in \texttt{scikit-learn} has shape \texttt{(n\_samples, n\_features)}, an array of tuples has shape \texttt{(n\_tuples, t, n\_features)}, where \texttt{t} is the number of elements in a tuple (e.g. 2 for pairs).
% Note that \texttt{scikit-learn} accepts arrays in a broad sense, in what they call "array-like" objects. \texttt{metric-learn} too accepts "tuple-like" objects.
Weakly supervised learning algorithms are similar to \texttt{scikit-learn} classifiers, but they \texttt{fit} on tuples that can be given as 3D arrays, instead of 2D arrays-like of points \texttt{X}. They have a \texttt{predict} method, that gives a prediction for each given tuple (e.g. for pairs it predicts whether the two points in a pair or similar or not). Tuples can be pairs or quadruplets, depending on the specific algorithm used. Below is a code example for pairs supervision:

\begin{minted}{python}
import numpy as np
from metric_learn import MMC
pairs = np.array([[[1.2, 3.2], [2.3, 5.5]], [[4.5, 2.3], [2.1, 2.3]]])
y_pairs = np.array([1, -1])
mmc = MMC()
mmc.fit(pairs, y_pairs)
\end{minted}
% mmc.predict(pairs)

This allows to use out of the box \texttt{scikit-learn}'s scoring functions, and therefore all routines for model selection, including the \texttt{GridSearchCV} object. Also note that for every weakly supervised learner, there is a \texttt{\_Supervised} version (e.g. for \texttt{MMC} there is \texttt{MMC\_Supervised}, that is a supervised version that samples pairs of similar samples from intra-class samples, and dissimilar samples from inter-class samples, and fits \texttt{MMC} on these pairs).

>>>>

\usepackage{jmlr2e}
\usepackage{minted}
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage{graphicx}
\usepackage{svg}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{stmaryrd}
\usepackage[colorinlistoftodos]{todonotes}
\newcommand{\aurelien}[1]{\todo[inline,caption={},color=orange!40]{{\it Aurelien:~}#1}}
\newcommand{\william}[1]{\todo[inline,caption={},color=blue!40]{{\it William:~}#1}}
%\renewcommand{\aurelien}[1]{}

\graphicspath{{images/}}

@@

\ShortHeadings{metric-learn: Metric Learning Algorithms in Python}{de Vazelhes, Carey, Tang, Vauquier and Bellet}
\firstpageno{1}

@@

\title{metric-learn: Metric Learning Algorithms in Python}

\author{\name William de Vazelhes \email william.de-vazelhes@inria.fr \\
       \addr INRIA, France
       \AND
       \name CJ Carey \email perimosocordiae@gmail.com \\
       \addr \aurelien{CJ must put in his affiliation}
       \AND
       \name Yuan Tang \email terrytangyuan@gmail.com \\
       \addr Ant Financial, United States
       \AND
       \name Nathalie Vauquier \email nathalie.vauquier@inria.fr \\
       \addr INRIA, France
       \AND
       \name Aur\'elien Bellet \email aurelien.bellet@inria.fr \\
       \addr INRIA, France
       }

@@

\begin{abstract}%   <- trailing '%' for backward compatibility of .sty file
\texttt{metric-learn} is an open source Python package implementing supervised and weakly-supervised distance metric learning algorithms. As part of \texttt{scikit-learn-contrib}, it provides a unified interface compatible with \texttt{scikit-learn} which allows to easily perform cross-validation, model selection, and pipelining with other machine learning estimators. \texttt{metric-learn} is thoroughly tested and available on PyPi under the MIT licence.
%The source code is available online at \url{http://github.com/scikit-learn-contrib/metric-learn}, as well as the documentation at \url{http://metric-learn.github.io/metric-learn/}.
\end{abstract}

@@

Many approaches in machine learning require a measure of distance between data
points. Traditionally, practitioners would choose a standard distance metric
(Euclidean, City-Block, Cosine, etc.) using a priori knowledge of the
domain. However, it is often difficult to design metrics that are well-suited
to the particular data and task of interest.
% \todo[inline]{probably need to add a sentence explaining what is meant by weak supervision}
Distance metric learning, or simply metric learning \citep{Bellet15,Kulis13}, aims at
automatically constructing task-specific distance metrics from data, in a machine learning manner. A key advantage of metric learning is that it can be applied beyond the standard supervised learning setting (data points associated with labels), in situations where only weaker forms of supervision are available (e.g., pairs of points that should be similar/dissimilar). The learned distance metric can
then be used to perform various tasks, for instance to retrieve the elements (images, documents) of a database that are semantically closer to a query element. It can also be plugged into other machine learning algorithms: popular use-cases include improving the accuracy of nearest neighbors learning models (for classification, regression, anomaly detection...) or biasing the clusters found by clustering algorithms like K-Means towards the intended semantics. Finally, metric learning can be used to perform dimensionality reduction.
These use-cases highlight the importance of integrating metric learning into the rest of the machine learning pipeline and tools.
% \todo[inline]{quick presentation of the package: implements several algorithms, with sklearn-compatible API for model selection, model evaluation, pipelining with other estimators. makes metric-learn stands out from other packages (maybe mention a few of them), in particular in the case of weakly supervised algorithms}

\texttt{metric-learn} is an open source package for metric learning in Python, which implements many popular metric-learning algorithms with different levels of supervision through a unified interface.
%\footnote{This makes \texttt{metric-learn} stand out from other packages, like \texttt{PyDML} (\url{https://github.com/jlsuarezdiaz/pyDML}) which focuses on supervised metric learning algorithms.}
% (not only class supervision, unlike most packages are focused on, including another package for metric learning in Python, \texttt{PyDML})
% , which is one interesting feature allowed by the metric learning paradigm. 
Its API is compatible with scikit-learn, the leading library for machine learning in Python, allowing for streamlined model selection, evaluation, and pipelining with other estimators.
The project is developed collaboratively through GitHub, its code is thoroughly tested using continuous integration tools. It was recently included to \texttt{scikit-learn-contrib}, which hosts high-quality \texttt{scikit-learn}-compatible projects.\footnote{\url{https://github.com/scikit-learn-contrib/scikit-learn-contrib}}
\aurelien{decide if we include or not a few sentences on positioning wrt other packages}

%, which makes \texttt{metric-learn} stand out from other packages, like \texttt{PyDML} \footnote{\url{https://github.com/jlsuarezdiaz/pyDML}}, which mostly focuses on supervised metric learning algorithms.
% Acknowledgements should go at the end, before appendices and references
%The rest of this paper is organized as follows. We first present an overview of the package, followed by a more detailed description of the available algorithms. We then give more details about the API and the software architecture, in particular the compatibility with \texttt{scikit-learn}, and conclude with a discussion of future developments.

% \aurelien{we could put the presentation of metric learning and supervision as Sec 2, and put the implemented algorithms in the overview of the package section as Section 3}

\section{Background on Metric Learning} \label{metriclearning}

% \todo[inline]{comparison with other packages ? }
% Metric learning problems fall into two main categories depending on the type
% of supervision available about the training data:

@@

% \paragraph{General setting} Metric learning problems are generally formulated as an optimization problem where one seeks to find, on the training data $(x_0, ..., x_n)$, the parameters $\theta$ of a distance function $D_\theta(x_i, x_j)$ that optimize some desired objective function $\mathcal{L}_\theta(x_0, ..., x_n) = f(D_\theta(x_0, x_0), ..., D_\theta(x_0, x_n), ..., D_\theta(x_n, x_0), ..., D_\theta(x_n, x_n))$, hoping that this metric will generalize the desired propertie(s) on new incoming test data.

Metric learning is generally formulated as an optimization problem where one seeks to find the parameters of a distance function that minimize some objective function over the input data.
All algorithms currently implemented in \texttt{metric-learn} learn so-called Mahalanobis distances. Given a real-valued parameter matrix $L$ of shape \texttt{(n\_components, n\_features)} where \texttt{n\_features} is the
number of features describing the data, the associated Mahalanobis distance between two points $x$ and $x'$ is defined as $D_L(x, x') = \sqrt{(Lx-Lx')^\top(Lx-Lx')}$.
This is equivalent to a Euclidean distance after the linear transformation of the feature space defined by $L$ (taking $L$ to be the identity matrix recovers the standard Euclidean distance).
Mahalanobis distance metric learning can thus be seen as learning a new
embedding space of (potentially reduced) dimension \texttt{n\_components}.
Note that $D_L$ can also be written as $D_L(x, x') = \sqrt{(x - x')^\top M (x - x')}$, where $M = L^\top L$ is called the Mahalanobis matrix.

@@

Metric learning algorithms can be classified according to the form of data supervision that they need to learn the metric. \texttt{metric-learn} currently implements algorithms that fall into the following categories:
\begin{itemize}
\item \textbf{Supervised learners} learn from a standard labeled dataset and essentially aim to bring closer together points from the same class while spreading further away points from a different class. For instance, the class may be the identity of a person in an image (see Figure~\ref{fig:full}). 
% \todo[inline]{ces figures ne sont pas très vendeuses. je propose de faire une figure conceptuelle en prenant l'exemple des faces expliquant la différence entre le supervisé, les paires et les quadruplets. on peut s'inspirer de \url{http://cedric.cnam.fr/~thomen/papers/Law_ICCV_2013_QWise.pdf}}
% \begin{figure}[H]
% \centering
% \begin{subfigure}{.5\textwidth}
%   \centering
%   \includegraphics[width=.8\linewidth]{supervised_without_metric.png}
%   \caption{Original points, before metric learning}
%   \label{fig:sub1}
% \end{subfigure}%
% \begin{subfigure}{.5\textwidth}
%   \centering
%   \includegraphics[width=.8\linewidth]{supervised_with_metric.png}
%   \caption{points after metric learning (ITML)}
%   \label{fig:sub2}
% \end{subfigure}
% \caption{Metric learning with pairwise constraints}
% \label{fig:test}
% \end{figure}
% \paragraph{Weakly Supervised Metric Learning algorithms}
% Weakly Supervised algorithms use less information than the previous algorithms: they only take tuple of points and possibly a corresponding label. See below for more concrete examples of algorithms learning on tuples.
\item \textbf{Pair learners.} 
Another way is to have pair of points as inputs, and a corresponding label indicating whether the two points are similar or not. We can then learn a metric that brings points from a similar pair closer together, and points from a dissimilar pair further away from each other. In many applications, this can be a much easier type of supervision to collect, in particular when there are many labels. %, since annotators will only need to say whether images are similar or not, and will not have to assign a label in a huge list of those.
For instance, it is much easier for an annotator to decide whether two images correspond to the same person than to associate an individual image to the right identity among a large number of people (see Figure~\ref{fig:pairs}). 
% \begin{figure}[H]
% \centering
% \begin{subfigure}{.5\textwidth}
%   \centering
%   \includegraphics[width=.8\linewidth]{pairs_without_metric.png}
%   \caption{Original points, before metric learning}
%   \label{fig:sub1}
% \end{subfigure}%
% \begin{subfigure}{.5\textwidth}
%   \centering
%   \includegraphics[width=.8\linewidth]{pairs_with_metric.png}
%   \caption{points after metric learning (ITML)}
%   \label{fig:sub2}
% \end{subfigure}
% \caption{Metric learning with pairwise constraints}
% \label{fig:test}
% \end{figure}
\item \textbf{Quadruplet Learners.}
The last possibility to learn a metric in \texttt{metric-learn} is to take as input quadruplets of points. We will learn a metric that brings the two first points of each quadruplet closer than the two last points. This can be used for instance to learn a metric space in which closer points have closer values of a given attributes. This is very useful if the attribute is not binary like in pair supervision (e.g. same person/not same person), but is rather continuous and difficult to annotate precisely (e.g., the age of a person on an image, see Figure \ref{fig:quadruplets}).
%for instance, we want to learn a metric that puts closer together people from approximately the same age.
\end{itemize}

\begin{figure}[H]
  \begin{subfigure}
     \includegraphics[scale=0.2]{labels.pdf}
     \label{fig:full}
\end{subfigure}
 \begin{subfigure}
     \includegraphics[scale=0.2]{pairs.pdf}
     \label{fig:pairs}
  \end{subfigure}
\begin{subfigure}
     \includegraphics[scale=0.2]{quadruplets.pdf}
     \label{fig:quadruplets}
  \end{subfigure}
  \caption{Different types of supervision for metric learning. (\subref{fig:full}) Full supervision: make samples from the same label (here, identity) closer together than samples from different labels. (\subref{fig:pairs}) Pair supervision: make similar pairs (here, corresponding to the same person) to be closer than dissimilar pairs (different persons). (\subref{fig:quadruplets}) Quadruplet supervision: make the two leftmost samples (here, people of similar age) to be closer than the two rightmost samples (of bigger age difference). (images are the funneled Labeled Faces in the Wild dataset \url{http://vis-www.cs.umass.edu/lfw/} \cite{Huang12}) (\aurelien{designing the 3 figures in inkscape to avoid the ugly tables would be better}}\aurelien{should also acknowledge dataset from which these images are taken}
  \label{fig:contour}
\end{figure}

\aurelien{update links, mention scikit-learn-contrib}

\texttt{metric-learn} only relies on core libraries from the SciPy ecosystem: \texttt{numpy}, \texttt{scipy}, and \texttt{scikit-learn}.
%\footnote{An optional dependency with \texttt{skggm} (\cite{skggm}) can be installed in order to use a more stable solver when using the LSML algorithm from the package.}
The current release (v.0.5.0) can be installed from the Python Package Index (PyPI), for python 2.7 and 3.5 or later. The source code is available on GitHub at \url{http://github.com/scikit-learn-contrib/metric-learn}. The development is collaborative and open to all contributors through the usual GitHub framework (issues and pull requests). The code is free of use, provided under the MIT licence. The interest of the package from the community is clear, since the project has been starred more than 740 times and forked more than 170 times at the time of writing.
A detailed documentation (including installation guidelines, the description of the algorithms and the API, as well as examples) is available at \url{http://contrib.scikit-learn.org/metric-learn}. The quality of the code is ensured by a thorough test coverage (96\% as of July 2019). Every new contribution to the code is automatically checked by a continuous integration platform that ensures a minimal test coverage of the added code, a minimal variation in test coverage of the overall project, and syntax formatting with \texttt{flake8}. %, which allows easy contribution for incoming developers, which can have their code automatically checked.
Currently, the following algorithms are implemented in \texttt{metric-learn}: for supervised metric learning: Neighborhood Components Analysis (NCA) (\cite{Goldberger04}), Large Margin Nearest Neighbors (LMNN) (\cite{Weinberger09}), Relative Components Analysis (RCA) (\cite{Shental02}), Local Fisher Discriminant Analysis (LFDA) (\cite{Sugiyama07}) and Metric Learning for Kernel Regression (MLKR) (\cite{Weinberger07}). Note that the latter is designed for regression problems and takes as input continuous labels. \footnote{\texttt{metric-learn} also provides an algorithm for learning on chunklets of data (groups of points from the same class)}\aurelien{maybe add footnote for RCA?}For pair metric learners: Mahalanobis Metric for Clustering (MMC) (\cite{Xiang08}), Sparse High-Dimensional Metric Learning (SDML) (\cite{Qi09}), Information Theoretic Metric Learning (ITML) (\cite{Davis07}). For learning on quadruplets: Metric Learning from Relative Comparisons by Minimizing Squared Residual (LSML) (\cite{Liu12}). The details description of these algorithms can be found in the sphinx documentation.

\aurelien{mention flake8 compliance and checks}

@@

\aurelien{mention optimization somewhere?}
\texttt{metric-learn} provides a unified interface to all metric learning algorithms (including the weakly supervised ones that have an input format that differs from the classic machine learning settings) and is designed to be fully compatible with the functionalities of \texttt{scikit-learn} \citep{scikit-learn}. All metric learners in \texttt{metric-learn} inherit from an abstract \texttt{BaseMetricLearner} object, which itself inherits from \texttt{BaseEstimator}. All classes inheriting from \texttt{BaseMetricLearner} should implement two methods, \texttt{get\_metric} and \texttt{score\_pairs}: the former returns a function that computes the distance between two points $u$ and $v$. The latter method computes the distances between a dataset of pairs of points.
\aurelien{score\_pairs and get\_metric should be mentioned here (all metric learners should implement these). In Mahalanobis, just mention they are implemented accordingly (do not delete: TODO: update accordingly mahalanobis}
% \todo[inline]{Should I talk about the classes etc ? Not useful since they are not visible by the user right ?}
% \todo[inline]{I think it can be nice, this is not documentation, but explaining how things are coded}

%\texttt{metric-learn}'s API is designed to be fully compatible with \texttt{scikit-learn} API (\cite{scikit-learn}).

@@

Supervised metric learners inherit from scikit-learn base class \texttt{TransformerMixin}, just like \texttt{scikit-learn}'s \texttt{LinearDiscriminantAnalysis} for instance. As such, they are compatible with pipelining (\texttt{sklearn.pipeline.Pipeline} and \texttt{sklearn.pipeline.make\_pipeline}.
% (they pass all scikit-learn's "estimators checks"). 
%  They can easily be pipelined with nearest neighbors estimators to improve their performance: \texttt{better\_knn = sklearn.pipeline.make\_pipeline(NCA(), KNeighborsClassifier())} (TODO: sth to explain NCA is in metric-learn)
\aurelien{mention pipelining}

\paragraph{Weakly Supervised metric learners} 
% Algorithms that learn on pairs or quadruplets take as input 3D array of shape \texttt{(n\_samples, t, n\_features)} (where...). In other words, sampling one element along the first dimension returns one tuple (e.g. a pair), that is a matrix of t lines corresponding to the t elements in the tuple.
% This format is what allows all these API choices, since slicing along the 1st dimension
% While a classical input array \texttt{X} in \texttt{scikit-learn} has shape \texttt{(n\_samples, n\_features)}, an array of tuples has shape \texttt{(n\_tuples, t, n\_features)}, where \texttt{t} is the number of elements in a tuple (e.g. 2 for pairs).
% Note that \texttt{scikit-learn} accepts arrays in a broad sense, in what they call "array-like" objects. \texttt{metric-learn} too accepts "tuple-like" objects.
Weakly supervised learning algorithms are similar to \texttt{scikit-learn} classifiers, but they \texttt{fit} on tuples that can be given as 3D arrays, instead of 2D arrays-like of points \texttt{X}. They have a \texttt{predict} method, that gives a prediction for each given tuple (e.g. for pairs it predicts whether the two points in a pair or similar or not). Tuples can be pairs or quadruplets, depending on the specific algorithm used. (see section~\ref{metriclearning}). Say we work with $n_t$ tuples of points with $p$ features each.\\
\textit{(i) Pairs learners} take as input an array-like \texttt{pairs} of shape $(n_t, 2, p)$, as well as an array-like \texttt{y\_pairs} of shape $(n_t,)$, which for each pair tells whether it contains similar or dissimilar samples. In order to \texttt{predict} on incoming pairs whether they are similar or not, they also need to fit a threshold, i.e. a reference distance with respect to which the distance between the incoming points will be compared. If this distance is higher than the threshold, the pair will be predicted as dissimilar, and if smaller as similar. This threshold is automatically fitted at fit time using a heuristic, but can be set manually using the method \texttt{set\_threshold}, or calibrated more specifically on a validation set, in order to optimize any classification score, using the method \texttt{calibrate\_threshold}.\\
\textit{(ii) Quadruplets learners} work on arrays-like of shape $(n_t, 4, p)$, and no labels array, where for each quadruplet the two first elements are the ones we want to be closer than the two last ones. They can also \texttt{predict} whether an incoming quadruplet is in the right order or not, i.e. whether the two first elements are indeed more similar that the two last, by comparing the two pairwise distances.
\aurelien{a bit more details would be welcome (explain both case of pairs and quadruplets). for pairs, need to mention threshold and calibration.}

\begin{minted}[%linenos=TRUE,
fontsize=\footnotesize]{python}
>>> from sklearn.model_selection import train_test_split, cross_val_score
>>> from sklearn.datasets import fetch_lfw_pairs
>>> from metric_learn import ITML  
>>> dataset = fetch_lfw_pairs(resize=0.2)  
>>> pairs, y_pairs = ((dataset.pairs.reshape(2200, 2, -1) - 127.5) / 255, 
                      2 * dataset.target - 1)
>>> pairs, _, y_pairs, _ = train_test_split(pairs, y_pairs, train_size=1000, 
                                            stratify=y_pairs, random_state=42) 
>>> itml = ITML(prior='random', random_state=42)  
>>> cross_val_score(itml, pairs, y_pairs, scoring='accuracy')
array([0.62275449, 0.65165165, 0.63963964])
\end{minted}
% mmc.predict(pairs)

\aurelien{the snippet does not work, cannot do 3-fold CV on 2 samples.. maybe we can load a subset of pairs from LFW? ideally we should also include the imports so the snippet is self-sufficient. to save space, maybe we can split this over 2 columns, with line numbering or something}
\william{new snippet takes large horizontal space, still split ?}
This allows to use out of the box \texttt{scikit-learn}'s scoring functions, and therefore all routines for model selection, including the \texttt{GridSearchCV} object. Also note that for every weakly supervised learner, there is a \texttt{\_Supervised} version (e.g. for \texttt{MMC} there is \texttt{MMC\_Supervised}, that is a supervised version that samples pairs of similar samples from intra-class samples, and dissimilar samples from inter-class samples, and fits \texttt{MMC} on these pairs).
